### 2018-02-13 paper.Rmd
Use sample data

---
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
library(ggplot2)
library(plyr)
library(sp)
library(rgdal)
library(ggalt)
library(dplyr)

long_norway <- c(0, 32)
lat_norway <- c(57, 72)

```

```{r exploration, include=FALSE, echo=FALSE}
source("exploration.R")
```

```{r author_info, include=FALSE, echo=FALSE}
first_name <- "Wouter"
last_name <- "Koch"
institution <- "Norwegian Biodiversity Information Centre"
address <- "Trondheim, Norway"
```

```{r read_ao_data, include=FALSE, echo=FALSE}
ao_data <- read_gbif(path="data/Laridae-NO-2016.csv")
```

# Norwegian citizen science data exploration

By `r last_name`, `r first_name`

> `r institution`, `r address`

---------------------------------------------

```{r filter_data, include=FALSE, echo=FALSE}
genera <- c("Larus")
ao_data_subset <- filter_by_genus(data=ao_data, genera=genera)

long_dataset <- c(min(ao_data_subset$decimallongitude) %/% 1 , max(ao_data_subset$decimallongitude) %/% 1 + 1)
lat_dataset <- c(min(ao_data_subset$decimallatitude) %/% 1 , max(ao_data_subset$decimallatitude) %/% 1 + 1)
# long_dataset <- long_norway
# lat_dataset <- lat_norway

```

A test data exploration of some [Norwegian Species Observation Service](https://www.artsobservasjoner.no) data, obtained through [GBIF](https://www.gbif.org) (N = `r nrow(ao_data)`).

Filtering on `r summarize_list(genera)` leaves `r nrow(ao_data_subset)` observations, plotted on the map below.

```{r map, echo=FALSE, message=FALSE}
world_map <- map_data("world", "Norway")

ggplot(ao_data_subset) + annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=decimallongitude, y=decimallatitude, color=species), alpha=0.3, size = .3) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
    xlab("Longitude") +
    ylab("Latitude")

```

Map for `r summarize_list(genera)` in Norway

```{r rasterize, message=FALSE, echo=FALSE}

  gridsizeinkm <- 50

  ao_data_subset <- add_rounded_coordinates(dataset=ao_data_subset, km=gridsizeinkm)
```

```{r rastermap, message=FALSE, echo=FALSE}
world_map <- map_data("world", "Norway")

ggplot(ao_data_subset) + 
    annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=species), alpha=0.3, size = gridsizeinkm * .02) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

Map for `r summarize_list(genera)` in Norway, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km  cells.

```{r densitymap, message=FALSE, echo=FALSE}

counts <- ddply(ao_data_subset, .(ao_data_subset$long_rounded, ao_data_subset$lat_rounded), nrow)
names(counts) <- c("long_rounded", "lat_rounded", "freq")

world_map <- map_data("world", "Norway")

ggplot(ao_data_subset, aes(decimallongitude, decimallatitude)) + 
    annotation_map(world_map, fill="white", color=NA) +
    geom_bin2d(aes(fill = log10(..count..)), binwidth = c(.018 * gridsizeinkm ,  .01  * gridsizeinkm)) +
    scale_fill_gradient(low = "gray90", high = "red") +
    annotation_map(world_map, fill=NA, color="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
      xlim=long_dataset,  # limits on longitude
      ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

Map for `r summarize_list(genera)` in Norway, raterized to `r .018 * gridsizeinkm` x `r .01  * gridsizeinkm` degree cells.

## Human population density

```{r human_pop_data, messages=FALSE, echo=FALSE}
geostat_data <- read_geostat(path="data/GeoStat_NO.csv")
geostat_data <- geostat_add_utm(geostat_data)

ggplot(geostat_data) + 
    annotation_map(world_map, fill="gray") +
    geom_point(aes(x=long, y=lat, color=log(X.TOT_P.)), size = .0004, shape=15) +
    scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

Log human population per square km in norway

```{r human_pop_rasterize, messages=FALSE, echo=FALSE}
geostat_data <- utm_add_rasterized(geostat_data, km=gridsizeinkm)
```

```{r human_pop_plot_rasterized, messages=FALSE, echo=FALSE}
ggplot(geostat_data) + 
    annotation_map(world_map, fill="gray") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(pop)), size = gridsizeinkm * .02) +
     scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

The log human population per `r gridsizeinkm` x `r gridsizeinkm` km cs cell in Norway.

```{r per_capita_scatterplot, messages=FALSE, echo=FALSE}
  num_vs_pop <- join_by_utm_rounded(ao_data_subset, geostat_data)

  ggplot(num_vs_pop, aes(x=pop, y=num)) + 
    geom_point()+
    geom_smooth(method=lm)
```

The number of `r summarize_list(genera)` in Norway per `r gridsizeinkm` x `r gridsizeinkm` km cs cells, versus the human population in those cells. R² of the linaer regression is `r summary(lm(num ~ pop, data=num_vs_pop))$r.squared`. R² of the log linaer regression is `r summary(lm(log10(num) ~ log10(pop), data=num_vs_pop))$r.squared`.

```{r per_capita_plot_rasterized, messages=FALSE, echo=FALSE}

num_vs_pop$percapita <- num_vs_pop$num / num_vs_pop$pop

ggplot(num_vs_pop) + 
    annotation_map(world_map, fill="gray") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(percapita)), size = gridsizeinkm * .02) +
     scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

The log number of `r summarize_list(genera)` observations in Norway per capita within `r gridsizeinkm` x `r gridsizeinkm` km cells.