### 2018-02-15 paper.Rmd
Generalize functions, uniform density maps

---
title: "Norwegian citizen science data exploration"
author: "Wouter Koch, NBIC"

output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
library(ggplot2)
library(plyr)
library(sp)
library(rgdal)
library(ggalt)
library(dplyr)

source("exploration.R")

obs_filter <- c("genus", "Larus")
```

******

```{r read_data, include=FALSE, echo=FALSE}
ao_data <- read_gbif(path="data/Laridae-NO-2016.csv")
geostat_data <- read_geostat(path="data/GeoStat_NO.csv")
```

```{r filter_data, include=FALSE, echo=FALSE}
ao_data_subset <- filter_observations(ao_data, obs_filter)  # filter observations

ao_data_subset <- rasterize(ao_data_subset)   # round all observation positions to km
geostat_data <- rasterize(geostat_data)       # round all population data positions to km

long_dataset <- c(min(ao_data_subset$decimallongitude) %/% 1 , max(ao_data_subset$decimallongitude) %/% 1 + 1)
lat_dataset <- c(min(ao_data_subset$decimallatitude) %/% 1 , max(ao_data_subset$decimallatitude) %/% 1 + 1)
```

A test data exploration of some [Norwegian Species Observation Service](https://www.artsobservasjoner.no) data, obtained through [GBIF](https://www.gbif.org) (N = `r nrow(ao_data)`).

Filtering on `r summarize_list(obs_filter[-1])` leaves `r nrow(ao_data_subset)` observations, plotted on the map below.

```{r map, echo=FALSE, message=FALSE}
world_map <- map_data("world", "Norway")

ggplot(ao_data_subset) + annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=decimallongitude, y=decimallatitude, color=species), alpha=0.5, size=cellsize/gridsizeinkm, shape=15) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
    xlab("Longitude") +
    ylab("Latitude")

```

Map for `r summarize_list(obs_filter[-1])` in Norway

```{r rasterize, message=FALSE, echo=FALSE}
  gridsizeinkm <- 50
  cellsize <- (lat_dataset[2] - lat_dataset[1]) * gridsizeinkm * .0016
  ao_data_subset <- rasterize(ao_data_subset, gridsizeinkm)
```

```{r rastermap, message=FALSE, echo=FALSE}
world_map <- map_data("world", "Norway")

ggplot(ao_data_subset) + 
    annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=species), alpha=0.3, size = .5 * cellsize) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

Map for `r summarize_list(obs_filter[-1])` in Norway, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km  cells.

```{r densitymap_2, message=FALSE, echo=FALSE}

counts <- ddply(ao_data_subset, .(ao_data_subset$long_rounded, ao_data_subset$lat_rounded), nrow)
names(counts) <- c("long_rounded", "lat_rounded", "freq")

ggplot(counts) + 
    annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(freq)), shape=15, size=cellsize) +
    scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")
```

Map for `r summarize_list(obs_filter[-1])` in Norway, raterized to `r .018 * gridsizeinkm` x `r .01  * gridsizeinkm` degree cells.

## Human population density

```{r human_pop_data, messages=FALSE, echo=FALSE}
ggplot(geostat_data) + 
    annotation_map(world_map, fill="gray") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(TOT_P)), size=cellsize/gridsizeinkm, shape=15) +
    scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

Log human population per square km in norway

```{r human_pop_rasterize, messages=FALSE, echo=FALSE}
geostat_data_rastered <- rasterize(geostat_data, gridsizeinkm)
```

```{r human_pop_plot_rasterized, messages=FALSE, echo=FALSE}

geostat_data_rastered <- geostat_data_rastered %>% group_by(utm_x_rounded, utm_y_rounded, lat_rounded, long_rounded) %>% summarise(TOT_P=sum(TOT_P))

ggplot(geostat_data_rastered) + 
    annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(TOT_P)), shape=15, size=cellsize) +
    scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

The log human population per `r gridsizeinkm` x `r gridsizeinkm` km cs cell in Norway.

```{r per_capita_scatterplot, messages=FALSE, echo=FALSE}
  num_vs_pop <- join_by_utm_rounded(ao_data_subset, geostat_data_rastered)

  ggplot(num_vs_pop, aes(x=TOT_P, y=num)) + 
    geom_point()+
    geom_smooth(method=lm)
```

The number of `r summarize_list(obs_filter[-1])` in Norway per `r gridsizeinkm` x `r gridsizeinkm` km cs cells, versus the human population in those cells. R² of the linaer regression is `r summary(lm(num ~ TOT_P, data=num_vs_pop))$r.squared`. R² of the log linaer regression is `r summary(lm(log10(num) ~ log10(TOT_P), data=num_vs_pop))$r.squared`.

```{r per_capita_plot_rasterized, messages=FALSE, echo=FALSE}

num_vs_pop$percapita <- num_vs_pop$num / num_vs_pop$TOT_P

ggplot(num_vs_pop) + 
    annotation_map(world_map, fill="gray50", colour="gray50") +
    geom_point(aes(x=long_rounded, y=lat_rounded, color=log(percapita)), shape=15, size=cellsize) +
    scale_color_distiller(palette = "Spectral") +
    annotation_map(world_map, fill=NA, colour="gray30") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5),
              xlim=long_dataset,  # limits on longitude
              ylim=lat_dataset) + # limits on latitude
   xlab("Longitude") + ylab("Latitude")

```

The log number of `r summarize_list(obs_filter[-1])` observations in Norway per capita within `r gridsizeinkm` x `r gridsizeinkm` km cells.