### 2018-04-16 absences.Rmd
Generate pseudoabsences

---
title: "Norwegian citizen science data - absence data"
author: "Wouter Koch, NBIC"

output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
  library(dplyr)
  library(rasterVis)
  library(rgbif)
  library(sp)
  library(raster)
  library(ggplot2)

  source("exploration.R")

  country_code <- "NO"
  country_coordinates <- c(3500000, 3500000, 5500000, 5500000)
  country_map <- map_data("world", country_code)

  ## Season onsets for Værnes airport according to Norwegian Meteorological Institute
  spring <- as.Date("2012-03-15",  format = "%Y-%m-%d") ## From Mar 15 
  summer <- as.Date("2012-05-22",  format = "%Y-%m-%d")  ## From May 22
  fall <- as.Date("2012-09-11",  format = "%Y-%m-%d")  ## From Sep 11
  winter <- as.Date("2012-11-18",  format = "%Y-%m-%d")  ## From Nov 18

  sp::coordinates(country_map) <- ~long + lat
  sp::proj4string(country_map) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
```

******

```{r variables, include=FALSE, echo=FALSE}
  higherLevel <- "Family"
  focalSpecies <- "Haliaeetus albicilla"

  gridsizeinkm <- 50
  grid <- get_grid(gridsizeinkm, country_coordinates)

```

```{r read_data, include=FALSE, echo=FALSE}
  higherLevelName <- name_backbone(name=focalSpecies, rank='species')[tolower(higherLevel)][[1]]
  taxon <- rgbif::name_suggest(q=higherLevelName, rank=higherLevel)

  gbif_filter <- list(
      paste('taxonKey =', taxon$key[1]),
      paste('country =', country_code),
      'datasetKey = b124e1e0-4755-430f-9eab-894f25a9b59c',
      'year = 2016',
      'hasCoordinate = TRUE',
      'hasGeospatialIssue = FALSE',
      type = "and")

  ao_data <- load_GBIF_data(gbif_filter)
  # geostat_data <- load_GeoStat_data(country_code)

```

```{r split_obs, include=FALSE, echo=FALSE}
  library(tidyr)

  ao_data <- subset(as.data.frame(ao_data), taxonRank == "SPECIES")
  ao_data <- ao_data %>% tidyr::separate_rows(recordedBy, sep = ",")
  ao_data$recordedBy <- as.factor(ao_data$recordedBy)

  per_user <- ao_data %>% group_by(recordedBy, species) %>% tally %>% tidyr::spread(species, n, fill = 0)
  per_user$observations <- (ao_data %>% group_by(recordedBy) %>% tally)$n

  per_user$species <- rowSums(per_user[c(-1)] > 0) - 1
  per_user <- per_user[c("recordedBy",focalSpecies,"observations","species")]

  names(per_user) <- c("recordedBy","observationsOfSpecies","totalObservations","numberOfSpeciesObserved")

```

```{r get_observers, include=FALSE, echo=FALSE}
  # Get all people that have reported this species, and at least x observations and y species
  per_user <- per_user %>% filter(observationsOfSpecies > 0 & totalObservations > 100 & numberOfSpeciesObserved > 3)

```

```{r get_empty_squares, include=FALSE, echo=FALSE}
  # For this species: get cells where it has been observed

  focus_species_observations <- ao_data %>% filter(species == focalSpecies)

  sp::coordinates(focus_species_observations) <- ~decimalLongitude + decimalLatitude
  sp::proj4string(focus_species_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
  focus_species_observations <- spTransform(focus_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
   focus_species_cells <- as.data.frame(focus_species_observations)
  focus_species_cells$decimalLatitude <- focus_species_cells$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)
  focus_species_cells$decimalLongitude <- focus_species_cells$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)

  focus_species_cells$cell <- paste(focus_species_cells$decimalLatitude, focus_species_cells$decimalLongitude, sep = "_")

  focus_species_cells <- focus_species_cells[c("cell")]
  focus_species_cells <- unique(focus_species_cells)

```

```{r get_absence, include=FALSE, echo=FALSE}
  # For such cells, find the number of filtered people that have been there
  # Or the number of species reported there by such filtered people

  other_species_observations <- ao_data %>% filter(species != focalSpecies)
  other_species_observations$date <- paste(other_species_observations$year, other_species_observations$month, other_species_observations$day, sep = "-")

  sp::coordinates(other_species_observations) <- ~decimalLongitude + decimalLatitude
  sp::proj4string(other_species_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
  other_species_observations <- spTransform(other_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))

  other_species_observations <- as.data.frame(other_species_observations)
  other_species_observations$decimalLatitude <- other_species_observations$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)
  other_species_observations$decimalLongitude <- other_species_observations$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)

  other_species_observations <- other_species_observations[c("recordedBy","decimalLatitude","decimalLongitude","date")]
  other_species_observations <- unique(other_species_observations)

  other_species_observations$cell <- paste(other_species_observations$decimalLatitude, other_species_observations$decimalLongitude, sep = "_")

  other_species_observations <- subset(other_species_observations, recordedBy %in% per_user$recordedBy)
  other_species_observations <- subset(other_species_observations, !(cell %in% focus_species_cells$cell))

  sp::coordinates(other_species_observations) <- ~decimalLongitude + decimalLatitude
  sp::proj4string(other_species_observations) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

  other_species_rasterized <- raster::rasterize(spTransform(other_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count')

  country_map <- spTransform(country_map, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))

```

```{r plots_presence, message=FALSE, echo=FALSE}

  focus_species_rasterized <- raster::rasterize(spTransform(focus_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count')

  gplot(focus_species_rasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="grey50", colour="black") +
    geom_tile(aes(fill=value), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    # scale_fill_continuous(high = "red", low = "white", na.value=NA) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Observations") + 
    ggtitle("Presence")
```

Observations of _`r focalSpecies`_ in Norway.

```{r plots_absence, message=FALSE, echo=FALSE}

  gplot(other_species_rasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="grey50", colour="black") +
    geom_tile(aes(fill=value), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    # scale_fill_continuous(high = "red", low = "white", na.value=NA) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Absences") + 
    ggtitle("Absence")

```

Absence of _`r focalSpecies`_ in Norway.
(Visits by users known to report _`r focalSpecies`_, where that user reported at least one observation of other species within the same `r tolower(higherLevel)` within that cell.)