### 2018-04-16 paper.Rmd
Generate pseudoabsences

---
title: "Norwegian citizen science data exploration"
author: "Wouter Koch, NBIC"

output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
  library(dplyr)
  library(rasterVis)
  library(rgbif)
  library(sp)
  library(raster)
  library(ggplot2)

  source("exploration.R")

  country_code <- "NO"
  country_coordinates <- c(3500000, 3500000, 5500000, 5500000)
  country_map <- map_data("world", country_code)

  ## Season onsets for Værnes airport according to Norwegian Meteorological Institute
  spring <- as.Date("2012-03-15",  format = "%Y-%m-%d") ## From Mar 15 
  summer <- as.Date("2012-05-22",  format = "%Y-%m-%d")  ## From May 22
  fall <- as.Date("2012-09-11",  format = "%Y-%m-%d")  ## From Sep 11
  winter <- as.Date("2012-11-18",  format = "%Y-%m-%d")  ## From Nov 18

  sp::coordinates(country_map) <- ~long + lat
  sp::proj4string(country_map) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
```

******

```{r read_data, include=FALSE, echo=FALSE}
  taxon <- rgbif::name_suggest(q='Larus', rank='Genus')

  gbif_filter <- list(
      paste('taxonKey =', taxon$key[1]),
      paste('country =', country_code),
      'datasetKey = b124e1e0-4755-430f-9eab-894f25a9b59c',
      'year = 2016',
      'hasCoordinate = TRUE',
      'hasGeospatialIssue = FALSE',
      type = "and")

  ao_data <- load_GBIF_data(gbif_filter)
  geostat_data <- load_GeoStat_data(country_code)

```

```{r grid, include=FALSE, echo=FALSE}
  gridsizeinkm <- 50

  grid <- get_grid(gridsizeinkm, country_coordinates)
  grid_10km <- get_grid(10, country_coordinates)

```

A test data exploration of some [Norwegian Species Observation Service](https://www.artsobservasjoner.no) data: observations within the `r tolower(taxon['rank'])` `r taxon['canonicalName']`, obtained through [GBIF](https://www.gbif.org) (N = `r nrow(ao_data)`).

## Observations

```{r map_per_species, echo=FALSE, message=FALSE}
  ggplot(subset(as.data.frame(ao_data),specificEpithet!="")) +
    annotation_map(as.data.frame(country_map), fill="white", colour="white") +
    geom_point(aes(x=decimalLongitude, y=decimalLatitude), color="darkgreen", alpha=.05, size=.25, shape=20) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "none") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    facet_wrap(~ species) + 
    ggtitle("Observations per species")

```

Map of observations of each species within the `r tolower(taxon['rank'])` `r taxon['canonicalName']` in Norway.

```{r map_per_season, echo=FALSE, message=FALSE}
  ggplot(as.data.frame(ao_data)) +
    annotation_map(as.data.frame(country_map), fill="white", colour="white") +
    geom_point(aes(x=decimalLongitude, y=decimalLatitude), color="darkgreen", alpha=.05, size=.25, shape=20) +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "none") +
    coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    facet_wrap(~ season) + 
    ggtitle("Observations per season")

```

Map of all observations within the `r tolower(taxon['rank'])` `r taxon['canonicalName']` in Norway, per season.

```{r rastermap_observations, message=FALSE, echo=FALSE}

  ao_data_rasterized <- raster::rasterize(spTransform(ao_data, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count')

  country_map <- spTransform(country_map, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))

  gplot(ao_data_rasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="grey50", colour="black") +
    geom_tile(aes(fill=log(value)), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    # scale_fill_continuous(high = "red", low = "white", na.value=NA) +
    scale_fill_distiller(palette = "Spectral", na.value=NA, direction = -1) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Observations (log)") + 
    ggtitle("Observation density")

```

Density map for all observations in Norway within the `r tolower(taxon['rank'])` `r taxon['canonicalName']`, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km cells.

```{r rastermap_observations_seasonal, message=FALSE, echo=FALSE}

  # ao_data_rasterized_winter <- raster::rasterize(spTransform(subset(ao_data, season=='Winter'), CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count', field='season')
  # ao_data_rasterized_spring <- raster::rasterize(spTransform(subset(ao_data, season=='Spring'), CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count', field='season')
  # ao_data_rasterized_summer <- raster::rasterize(spTransform(subset(ao_data, season=='Summer'), CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count', field='season')
  # ao_data_rasterized_fall <- raster::rasterize(spTransform(subset(ao_data, season=='Fall'), CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count', field='season')

  # ao_season_stack <- stack(ao_data_rasterized_winter, ao_data_rasterized_spring, ao_data_rasterized_summer, ao_data_rasterized_fall)
  # ao_season_stack <- log(ao_season_stack)
  # names(ao_season_stack) <- c('Winter','Spring','Summer','Fall')
  # 
  # levelplot(ao_season_stack,
  #     col.regions=colorRampPalette(rev(brewer.pal(5,"Spectral"))),
  #     scales=list(draw=FALSE ))

```

```{r rastermap_observations_seasonal_exp, message=FALSE, echo=FALSE}
#   
# 
#   
#   
  # ao_data_rasterized_winter <- ratify(ao_data_rasterized_winter)
  # rat <- levels(ao_data_rasterized_winter)[[1]]
  # rat$season <- "Winter"
  # levels(ao_data_rasterized_winter) <- rat
  # 
  # ao_data_rasterized_spring <- ratify(ao_data_rasterized_spring)
  # rat <- levels(ao_data_rasterized_spring)[[1]]
  # rat$season <- "Spring"
  # levels(ao_data_rasterized_spring) <- rat
  # 
  # ao_data_rasterized_summer <- ratify(ao_data_rasterized_summer)
  # rat <- levels(ao_data_rasterized_summer)[[1]]
  # rat$season <- "Summer"
  # levels(ao_data_rasterized_summer) <- rat
  # 
  # ao_data_rasterized_fall <- ratify(ao_data_rasterized_fall)
  # rat <- levels(ao_data_rasterized_fall)[[1]]
  # rat$season <- "Fall"
  # levels(ao_data_rasterized_fall) <- rat
  # 
  # 
  # ao_season_stack <- stack(ao_data_rasterized_winter, ao_data_rasterized_spring, ao_data_rasterized_summer, ao_data_rasterized_fall)

#   
#   
#   
   # View(as.data.frame(ao_data_rasterized_winter))
#   
# 
  # gplot(ao_data_rasterized_winter) +
  #   geom_tile(aes(fill = log(value))) +
  #   facet_wrap(~ name) +
  #   scale_fill_gradient(low = 'white', high = 'blue') +
  #   coord_equal()
# 
# 
#   ao_data_rasterized_hm<- raster::rasterize(spTransform(ao_data, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count', field='season', na.rm=TRUE)
#     
#     
#     
#     
# 
  # country_map <- spTransform(country_map, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
#   
# gplot(ao_season_stack) +
#     coord_fixed() +
#     annotation_map(as.data.frame(country_map), fill="gray40", colour="black") +
#     geom_tile(aes(fill=log(value)), alpha=1) +
#     annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
#     theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
#     scale_fill_distiller(palette = "Spectral", na.value=NA) +
#     guides(color = guide_legend(override.aes = list(size=4))) +
#     facet_wrap(~ layer) +
#     # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
#     xlab("Longitude") +
#     ylab("Latitude")

```

## Human population

```{r map_human_population, messages=FALSE, echo=FALSE}
  pop_unrasterized <- raster::rasterize(geostat_data, grid_10km, "TOT_P", fun='sum')

  gplot(pop_unrasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="gray50", colour="black") +
    geom_tile(aes(fill=log(value)), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    scale_fill_distiller(palette = "Spectral", na.value=NA, direction=-1) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Inhabitants (log)") +
    ggtitle("Human population density")
```

Log human population per 10 x 10 km cell in norway

```{r rastermap_human_population, messages=FALSE, echo=FALSE}

pop_rasterized <- raster::rasterize(geostat_data, grid, "TOT_P", fun='sum')

gplot(pop_rasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="gray50", colour="black") +
    geom_tile(aes(fill=log(value)), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    scale_fill_distiller(palette = "Spectral", na.value=NA, direction=-1) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Inhabitants (log)") +
    ggtitle("Human population density")

```

The log human population in Norway, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km cells.

```{r scatterplot_per_capita, messages=FALSE, echo=FALSE}

  pop_rasterized.points <- as.data.frame(rasterToPoints(pop_rasterized, spatial=TRUE))[c("layer","x","y")]
  pop_rasterized.points <- plyr::rename(pop_rasterized.points, c("layer"="population"))
  ao_data_rasterized.points <- as.data.frame(rasterToPoints(ao_data_rasterized, spatial=TRUE))[c("ID","x","y")]
  ao_data_rasterized.points <- plyr::rename(ao_data_rasterized.points, c("ID"="observations"))

  num_vs_pop <- ao_data_rasterized.points %>% inner_join(pop_rasterized.points, by = c("x", "y"))

  # num_vs_pop <- subset(num_vs_pop, population<100000 & observations < 1000)

  sp::coordinates(num_vs_pop) <- ~x + y
  sp::proj4string(num_vs_pop) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

  ggplot(as.data.frame(num_vs_pop), aes(x=population, y=observations)) +
    geom_point()+
    geom_smooth(method=lm)
```

The number of observations in Norway within the `r tolower(taxon['rank'])` `r taxon['canonicalName']` per `r gridsizeinkm` x `r gridsizeinkm` km cells, versus the human population in those cells. R² of the linear regression is `r summary(lm(observations ~ population, data=num_vs_pop))$r.squared`.

```{r rastermap_observations_per_capita, messages=FALSE, echo=FALSE}

num_vs_pop$percapita <- num_vs_pop$observations / num_vs_pop$population
num_vs_pop_rasterized <- raster::rasterize(num_vs_pop, grid, "percapita", fun='sum')

gplot(num_vs_pop_rasterized) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="gray50", colour="black") +
    geom_tile(aes(fill=log(value)), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    scale_fill_distiller(palette = "Spectral", na.value=NA) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Observations (log)") +
    ggtitle("Observations per capita")
```

The log number of observations in Norway within the `r tolower(taxon['rank'])` `r taxon['canonicalName']`, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km cells.

```{r rastermap_observations_relative_to_lm, messages=FALSE, echo=FALSE}

num_vs_pop <- as.data.frame(num_vs_pop)

B0 <- (summary(lm(observations ~ population, data=num_vs_pop))$coefficients)["(Intercept)","Estimate"]
B1 <- (summary(lm(observations ~ population, data=num_vs_pop))$coefficients)["population","Estimate"]

num_vs_pop$relativeToLM <- num_vs_pop$observations / (B0 + B1 * num_vs_pop$population)

sp::coordinates(num_vs_pop) <- ~x + y
sp::proj4string(num_vs_pop) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

num_vs_pop_rasterized_LM <- raster::rasterize(num_vs_pop, grid, "relativeToLM", fun='sum')

gplot(num_vs_pop_rasterized_LM) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="gray50", colour="black") +
    geom_tile(aes(fill=log(value)), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    scale_fill_distiller(palette = "Spectral", na.value=NA) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "Observations (log)") +
    ggtitle("Observations corrected for inhabitants")
```

The log number of observations within the `r tolower(taxon['rank'])` `r taxon['canonicalName']` in Norway, divided by the number predicted by the regression line (numberOfObservations = `r B0` + `r B1` * humanPopulation), rasterized to `r gridsizeinkm` x `r gridsizeinkm` km cells.

## Bioclimatic variables

```{r climate_data, messages=FALSE, echo=FALSE}

env <- getData('worldclim', var='bio', res=5)

env2 <- projectRaster(env[[1]], grid, method="bilinear")

# plot(env, 1, main=NULL, axes=FALSE)

gplot(env2) +
    coord_fixed() +
    annotation_map(as.data.frame(country_map), fill="gray50", colour="black") +
    geom_tile(aes(fill=value/10), alpha=1) +
    annotation_map(as.data.frame(country_map), fill=NA, colour="black") +
    theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") +
    scale_fill_distiller(palette = "Spectral", na.value=NA, direction=-1) +
    guides(color = guide_legend(override.aes = list(size=4))) +
    # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill = "°C") +
    ggtitle("Annual Mean Temperature")

```

The Annual Mean Temperature in Norway, rasterized to `r gridsizeinkm` x `r gridsizeinkm` km cells.