### 2018-05-09 absences.Rmd
Plot species and observations per pseudoevent

---
title: "Norwegian citizen science data - absence data"
author: "Wouter Koch, NBIC"

output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
  library(dplyr)
  library(rasterVis)
  library(rgbif)
  library(sp)
  library(raster)
  library(ggplot2)

  source("exploration.R")

  country_code <- "NO"
  country_coordinates <- c(3500000, 3500000, 5500000, 5500000)
  country_map <- map_data("world", country_code)

  ## Season onsets for Værnes airport according to Norwegian Meteorological Institute
  spring <- as.Date("2012-03-15",  format = "%Y-%m-%d") ## From Mar 15 
  summer <- as.Date("2012-05-22",  format = "%Y-%m-%d")  ## From May 22
  fall <- as.Date("2012-09-11",  format = "%Y-%m-%d")  ## From Sep 11
  winter <- as.Date("2012-11-18",  format = "%Y-%m-%d")  ## From Nov 18

  sp::coordinates(country_map) <- ~long + lat
  sp::proj4string(country_map) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
  country_map <- spTransform(country_map, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
```

******

```{r variables, include=FALSE, echo=FALSE}
  higherLevel <- "Class"
  focalSpecies <- "Haliaeetus albicilla"

  gridsizeinkm <- 20
  grid <- get_grid(gridsizeinkm, country_coordinates)

```

```{r read_data, include=FALSE, echo=FALSE}
  higherLevelName <- name_backbone(name=focalSpecies, rank='species')[tolower(higherLevel)][[1]]
  taxon <- rgbif::name_suggest(q=higherLevelName, rank=higherLevel)

  gbif_filter <- list(
      paste('taxonKey =', taxon$key[1]),
      paste('country =', country_code),
      'datasetKey = b124e1e0-4755-430f-9eab-894f25a9b59c',
      'year = 2016',
      'hasCoordinate = TRUE',
      'hasGeospatialIssue = FALSE',
      type = "and")

  ao_data <- load_GBIF_data(gbif_filter)
  # geostat_data <- load_GeoStat_data(country_code)

```

```{r split_obs, include=FALSE, echo=FALSE}
  library(tidyr)

  ao_data <- subset(as.data.frame(ao_data), taxonRank == "SPECIES")
  ao_data <- ao_data %>% tidyr::separate_rows(recordedBy, sep = ",")
  ao_data$recordedBy <- as.factor(ao_data$recordedBy)

  # per_user <- ao_data %>% group_by(recordedBy, species) %>% tally %>% tidyr::spread(species, n, fill = 0)
  # per_user$observations <- (ao_data %>% group_by(recordedBy) %>% tally)$n
  # 
  # per_user$species <- rowSums(per_user[c(-1)] > 0) - 1
  # per_user <- per_user[c("recordedBy",focalSpecies,"observations","species")]
  # 
  # names(per_user) <- c("recordedBy","observationsOfSpecies","totalObservations","numberOfSpeciesObserved")

```

```{r ao_data_to_wide, include=FALSE, echo=FALSE}

ao_data_wide <- ao_data[c("decimalLatitude","decimalLongitude","year","month","day","recordedBy","species")]

ao_data_wide <- spread(count(ao_data_wide, decimalLatitude, decimalLongitude, year, month, day, recordedBy, species), species, n, fill = 0)

```

```{r get_sums, message=FALSE, echo=FALSE}
counted_species <- rowSums(ao_data_wide[,(7:ncol(ao_data_wide))] > 0)
counted_observations <- rowSums(ao_data_wide[,(7:ncol(ao_data_wide))])

```

```{r plot_species_counts, message=FALSE, echo=FALSE}
ao_data_wide$species <- counted_species
species_counts <- table(ao_data_wide$species)
barplot(species_counts, main="Number of species per pseudoevent", xlab="Number of species")
```

```{r plot_observation_counts, message=FALSE, echo=FALSE}
ao_data_wide$observations <- counted_observations
observation_counts <- table(ao_data_wide$observations)
barplot(observation_counts, main="Number of observations per pseudoevent", xlab="Number of observations")
```

```{r snap_observations_to_raster, include=FALSE, echo=FALSE}

ao_data_snapped_to_grid <- ao_data

sp::coordinates(ao_data_snapped_to_grid) <- ~decimalLongitude + decimalLatitude
sp::proj4string(ao_data_snapped_to_grid) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
ao_data_snapped_to_grid <- spTransform(ao_data_snapped_to_grid, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"))
ao_data_snapped_to_grid <- as.data.frame(ao_data_snapped_to_grid)
ao_data_snapped_to_grid$decimalLatitude <- ao_data_snapped_to_grid$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)
ao_data_snapped_to_grid$decimalLongitude <- ao_data_snapped_to_grid$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500)

```

```{r snap_observations_to_raster, include=FALSE, echo=FALSE}

ao_data_snapped_to_grid_wide <- ao_data_snapped_to_grid[c("decimalLatitude","decimalLongitude","year","month","day","recordedBy","species")]

ao_data_snapped_to_grid_wide <- spread(count(ao_data_snapped_to_grid_wide, decimalLatitude, decimalLongitude, year, month, day, recordedBy, species), species, n, fill = 0)

```

```{r get_sums_snapped, message=FALSE, echo=FALSE}
counted_species_snapped <- rowSums(ao_data_snapped_to_grid_wide[,(7:ncol(ao_data_snapped_to_grid_wide))] > 0)

counted_observations_snapped <- rowSums(ao_data_snapped_to_grid_wide[,(7:ncol(ao_data_snapped_to_grid_wide))])

```

```{r plot_snapped_species_counts, message=FALSE, echo=FALSE}
ao_data_snapped_to_grid_wide$species <- counted_species_snapped
species_counts_snapped <- table(ao_data_snapped_to_grid_wide$species)
barplot(species_counts_snapped, main="Number of species per pseudoevent, snapped to grid", xlab="Number of species")
```

```{r plot_snapped_observation_counts, message=FALSE, echo=FALSE}
ao_data_snapped_to_grid_wide$observations <-counted_observations_snapped
observation_counts_snapped <- table(ao_data_snapped_to_grid_wide$observations)
barplot(observation_counts_snapped, main="Number of observations per pseudoevent, snapped to grid", xlab="Number of observations")
```

<!-- ```{r get_observers, include=FALSE, echo=FALSE} -->
<!--   # Get all people that have reported this species, and at least x observations and y species -->
<!--   per_user <- per_user %>% filter(observationsOfSpecies > 0 & totalObservations > 100 & numberOfSpeciesObserved > 3) -->

<!-- ``` -->

<!-- ```{r get_presence_observations , include=FALSE, echo=FALSE} -->
<!--   # Get one observation per observer per day per rastercell of the focal species -->
<!--   presence_observations <- ao_data %>% filter(species == focalSpecies) -->
<!--   presence_observations <- subset(presence_observations, recordedBy %in% per_user$recordedBy) -->

<!--   sp::coordinates(presence_observations) <- ~decimalLongitude + decimalLatitude -->
<!--   sp::proj4string(presence_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0") -->
<!--   presence_observations <- spTransform(presence_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")) -->
<!--   presences <- as.data.frame(presence_observations) -->
<!--   presences$decimalLatitude <- presences$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   presences$decimalLongitude <- presences$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->

<!--   presences$cell <- paste(presences$decimalLatitude, presences$decimalLongitude, sep = "_") -->

<!--   presences <- presences[c("decimalLatitude","decimalLongitude","year","month","day","recordedBy")] -->
<!--   presences <- unique(presences) -->
<!--   presences$summary <- paste( -->
<!--                                             presences$decimalLatitude, -->
<!--                                             presences$decimalLongitude, -->
<!--                                             presences$year, -->
<!--                                             presences$month, -->
<!--                                             presences$day, -->
<!--                                             presences$recordedBy, -->
<!--                                             sep = "_") -->

<!-- ``` -->

<!-- ```{r get_other_observations , include=FALSE, echo=FALSE} -->
<!--   # Get one observation per observer per day per rastercell of the non-focal species -->
<!--   other_observations <- ao_data %>% filter(species != focalSpecies) -->
<!--   other_observations <- subset(other_observations, recordedBy %in% per_user$recordedBy) -->

<!--   sp::coordinates(other_observations) <- ~decimalLongitude + decimalLatitude -->
<!--   sp::proj4string(other_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0") -->
<!--   other_observations <- spTransform(other_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")) -->
<!--   other_observations_cells <- as.data.frame(other_observations) -->
<!--   other_observations_cells$decimalLatitude <- other_observations_cells$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   other_observations_cells$decimalLongitude <- other_observations_cells$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->

<!--   other_observations_cells$cell <- paste(other_observations_cells$decimalLatitude, other_observations_cells$decimalLongitude, sep = "_") -->

<!--   other_observations_cells <- other_observations_cells[c("decimalLatitude","decimalLongitude","year","month","day","recordedBy")] -->
<!--   other_observations_cells <- unique(other_observations_cells) -->
<!--   other_observations_cells$summary <- paste( -->
<!--                                             other_observations_cells$decimalLatitude, -->
<!--                                             other_observations_cells$decimalLongitude, -->
<!--                                             other_observations_cells$year, -->
<!--                                             other_observations_cells$month, -->
<!--                                             other_observations_cells$day, -->
<!--                                             other_observations_cells$recordedBy, -->
<!--                                             sep = "_") -->

<!-- ``` -->

<!-- ```{r get_absences , include=FALSE, echo=FALSE} -->
<!--   # Get absences -->
<!--   absences <- subset(other_observations_cells, !(summary %in% presences$summary)) -->

<!--   sp::coordinates(absences) <- ~decimalLongitude + decimalLatitude -->
<!--   sp::proj4string(absences) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs") -->

<!--   sp::coordinates(presences) <- ~decimalLongitude + decimalLatitude -->
<!--   sp::proj4string(presences) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs") -->

<!-- ``` -->

<!-- ```{r plots_absence, message=FALSE, echo=FALSE} -->

<!--   absences_rasterized <- raster::rasterize(spTransform(absences, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count') -->

<!--   gplot(absences_rasterized) + -->
<!--     coord_fixed() + -->
<!--     annotation_map(as.data.frame(country_map), fill="white", colour="grey50") + -->
<!--     geom_tile(aes(fill=log(value)), alpha=1) + -->
<!--     scale_fill_gradient(low = "yellow", high = "red", na.value = NA) + -->
<!--     theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") + -->
<!--     guides(color = guide_legend(override.aes = list(size=4))) + -->
<!-- #    annotation_map(as.data.frame(country_map), fill=NA, colour="black") + -->
<!--     xlab("Longitude") + -->
<!--     ylab("Latitude") + -->
<!--     labs(fill = "Absences") + -->
<!--     ggtitle("Absence") -->

<!-- ``` -->

<!-- ```{r plots_presence, message=FALSE, echo=FALSE} -->
<!--   presences_rasterized <- raster::rasterize(spTransform(presences, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count') -->

<!--   gplot(presences_rasterized) + -->
<!--     coord_fixed() + -->
<!--     annotation_map(as.data.frame(country_map), fill="white", colour="grey50") + -->
<!--     geom_tile(aes(fill=log(value)), alpha=1) + -->
<!--     scale_fill_gradient(low = "yellow", high = "darkblue", na.value = NA) + -->
<!-- #    annotation_map(as.data.frame(country_map), fill=NA, colour="black") + -->
<!--     theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") + -->
<!--     guides(color = guide_legend(override.aes = list(size=4))) + -->
<!--     xlab("Longitude") + -->
<!--     ylab("Latitude") + -->
<!--     labs(fill = "Presences") + -->
<!--     ggtitle("Presence") -->

<!-- ``` -->

<!-- ```{r plots_presence, message=FALSE, echo=FALSE} -->

<!--   presence_minus_absence <- overlay(presences_rasterized, absences_rasterized, fun=function(r1, r2){return(r1/r2)}) -->

<!--   gplot(presence_minus_absence) + -->
<!--     coord_fixed() + -->
<!--     annotation_map(as.data.frame(country_map), fill="white", colour="grey50") + -->
<!--     geom_tile(aes(fill=log(value)), alpha=1) + -->
<!--     scale_fill_gradient2(low = "red", high = "darkblue", mid = "yellow", na.value = NA) + -->
<!--     theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") + -->
<!--     guides(color = guide_legend(override.aes = list(size=4))) + -->
<!--     xlab("Longitude") + -->
<!--     ylab("Latitude") + -->
<!--     labs(fill = "log(Presences / absences)") + -->
<!--     ggtitle("Relative absence") -->

<!-- ``` -->

<!-- ```{r get_empty_squares, include=FALSE, echo=FALSE} -->
<!--   # For this species: get cells where it has been observed -->

<!--   # focus_species_observations <- ao_data %>% filter(species == focalSpecies) -->
<!--   #  -->
<!--   # sp::coordinates(focus_species_observations) <- ~decimalLongitude + decimalLatitude -->
<!--   # sp::proj4string(focus_species_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0") -->
<!--   # focus_species_observations <- spTransform(focus_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")) -->
<!--   #  focus_species_cells <- as.data.frame(focus_species_observations) -->
<!--   # focus_species_cells$decimalLatitude <- focus_species_cells$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   # focus_species_cells$decimalLongitude <- focus_species_cells$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   #  -->
<!--   # focus_species_cells$cell <- paste(focus_species_cells$decimalLatitude, focus_species_cells$decimalLongitude, sep = "_") -->
<!--   #  -->
<!--   # focus_species_cells <- focus_species_cells[c("cell")] -->
<!--   # focus_species_cells <- unique(focus_species_cells) -->

<!-- ``` -->

<!-- ```{r get_absence, include=FALSE, echo=FALSE} -->
<!--   # For such cells, find the number of filtered people that have been there -->
<!--   # Or the number of species reported there by such filtered people -->

<!--   # other_species_observations <- ao_data %>% filter(species != focalSpecies) -->
<!--   # other_species_observations$date <- paste(other_species_observations$year, other_species_observations$month, other_species_observations$day, sep = "-") -->
<!--   #  -->
<!--   # sp::coordinates(other_species_observations) <- ~decimalLongitude + decimalLatitude -->
<!--   # sp::proj4string(other_species_observations) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0") -->
<!--   # other_species_observations <- spTransform(other_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")) -->
<!--   #  -->
<!--   # other_species_observations <- as.data.frame(other_species_observations) -->
<!--   # other_species_observations$decimalLatitude <- other_species_observations$decimalLatitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   # other_species_observations$decimalLongitude <- other_species_observations$decimalLongitude %/% (gridsizeinkm * 1000) * (gridsizeinkm * 1000) + (gridsizeinkm * 500) -->
<!--   #  -->
<!--   # other_species_observations <- other_species_observations[c("recordedBy","decimalLatitude","decimalLongitude","date")] -->
<!--   # other_species_observations <- unique(other_species_observations) -->
<!--   #  -->
<!--   # other_species_observations$cell <- paste(other_species_observations$decimalLatitude, other_species_observations$decimalLongitude, sep = "_") -->
<!--   #  -->
<!--   # other_species_observations <- subset(other_species_observations, recordedBy %in% per_user$recordedBy) -->
<!--   # other_species_observations <- subset(other_species_observations, !(cell %in% focus_species_cells$cell)) -->
<!--   #  -->
<!--   #  -->
<!--   # sp::coordinates(other_species_observations) <- ~decimalLongitude + decimalLatitude -->
<!--   # sp::proj4string(other_species_observations) <- CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs") -->
<!--   #  -->
<!--   # other_species_rasterized <- raster::rasterize(spTransform(other_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count') -->
<!--   #  -->
<!--   #  -->
<!--   # country_map <- spTransform(country_map, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")) -->

<!-- ```   -->

<!-- ```{r plots_presence, message=FALSE, echo=FALSE} -->

<!--   # focus_species_rasterized <- raster::rasterize(spTransform(focus_species_observations, CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")), grid, fun='count') -->
<!--   #  -->
<!--   # gplot(focus_species_rasterized) + -->
<!--   #   coord_fixed() + -->
<!--   #   annotation_map(as.data.frame(country_map), fill="grey50", colour="black") + -->
<!--   #   geom_tile(aes(fill=value), alpha=1) + -->
<!--   #   annotation_map(as.data.frame(country_map), fill=NA, colour="black") + -->
<!--   #   theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") + -->
<!--   #   # scale_fill_continuous(high = "red", low = "white", na.value=NA) + -->
<!--   #   guides(color = guide_legend(override.aes = list(size=4))) + -->
<!--   #   # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) + -->
<!--   #   xlab("Longitude") + -->
<!--   #   ylab("Latitude") + -->
<!--   #   labs(fill = "Observations") +  -->
<!--   #   ggtitle("Presence") -->
<!-- ``` -->

<!-- Observations of _`r focalSpecies`_ in Norway. -->

<!-- ```{r plots_absence, message=FALSE, echo=FALSE} -->

<!--   # gplot(other_species_rasterized) + -->
<!--   #   coord_fixed() + -->
<!--   #   annotation_map(as.data.frame(country_map), fill="grey50", colour="black") + -->
<!--   #   geom_tile(aes(fill=value), alpha=1) + -->
<!--   #   annotation_map(as.data.frame(country_map), fill=NA, colour="black") + -->
<!--   #   theme(panel.background = element_rect(fill="aliceblue"), legend.position = "right") + -->
<!--   #   # scale_fill_continuous(high = "red", low = "white", na.value=NA) + -->
<!--   #   guides(color = guide_legend(override.aes = list(size=4))) + -->
<!--   #   # coord_map(projection = "gnomonic", orientation=c(60, 16, -5)) + -->
<!--   #   xlab("Longitude") + -->
<!--   #   ylab("Latitude") + -->
<!--   #   labs(fill = "Absences") +  -->
<!--   #   ggtitle("Absence") -->

<!-- ``` -->

<!-- Absence of _`r focalSpecies`_ in Norway. -->
<!-- (Visits by users known to report _`r focalSpecies`_, where that user reported at least one observation of other species within the same `r tolower(higherLevel)` within that cell.) -->